pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single cell RNA-seq preprocessing tools

### 2020-07-03 bmc_article.tex
Updates from Overleaf

\parttitle{Background} The massive growth of single-cell RNA-sequencing (scRNAseq) and the methods for its analysis still lack sufficient and up-to-date benchmarks that could guide analytical choices. Numerous benchmark studies already exist and cover most of scRNAseq processing and analytical methods but only a few give advice on a comprehensive pipeline. Moreover, current studies often focused on isolated steps of the process and do not address the impact of a tool on both the intermediate and the final steps of the analysis.

@@

The evaluation of \textit{scVI} algorithm was performed separate to the other pipeline combinations since the model training on our reference datasets was especially resource-consuming {\color{red} and follows a different workflow from the other pipelines. In particular, for datasets with fewer cells than genes, \textit{scVI} requires a feature selection to be performed prior to the normalization and imputation steps; furthermore, its dimensional reduction steps rely directly on raw counts. We therefore evaluated it separately and compared it to a representative subset of pipelines using \textit{sctransform} normalization and \textit{Seurat}'s PCA dimension reduction method. The resulting evaluation is available in the Supplementary Figure 13.

Using \textit{scVI}, we considered as imputed values the mean of the negative binomial distribution modeling the counts (accessed through the \textit{imputation} method). Following \textit{scVI}'s online documentation, we used the \textit{VAE} model, using a train size of 1, a training on 400 epochs and a learning rate of 0.001. As a major limitation of \textit{scVI} is the training on dataset with low gene to cell ratio, the features were subsampled to 2000 using the \textit{subsample\_genes} function, which resembles the \textit{Seurat} V3 VST gene selection method. All other parameters were left unmodified. The normalized values were computed from the normalized mean of the negative binomial distribution modeling the counts (accessed through the \textit{get\_sample\_scale} method).

>>>>

\parttitle{Background} Despite the massive growth of single-cell RNA-sequencing (scRNAseq) and corresponding computational methods, sufficient and up-to-date benchmarks to guide analytical choices are still lacking. Numerous benchmark studies already exist and cover most of scRNAseq processing and analytical methods but only a few give advice on a comprehensive pipeline. Moreover, current studies often focused on isolated steps of the process and do not address the impact of a tool on both the intermediate and the final steps of the analysis.

@@

The evaluation of \textit{scVI} algorithm was performed separate to the other pipeline combinations since the model training on our reference datasets was especially resource-consuming {\color{red} and follows a different workflow from the other pipelines. In particular, for datasets with fewer cells than genes, \textit{scVI} requires a feature selection to be performed prior to the normalization step; furthermore, its dimensional reduction steps rely directly on raw counts. We therefore evaluated it separately and compared it to a representative subset of pipelines using \textit{sctransform} normalization and \textit{Seurat}'s PCA dimension reduction method. The resulting evaluation is available in the Supplementary Figure 13.

Using \textit{scVI}, we considered as normalized values the normalized mean of the negative binomial distribution modeling the counts (accessed through the \textit{get\_sample\_scale} method). Following \textit{scVI}'s online documentation, we used the \textit{VAE} model, using a train size of 1, a training on 400 epochs and a learning rate of 0.001. As a major limitation of \textit{scVI} is the training on dataset with low gene to cell ratio, the features were subsampled to 2000 using the \textit{subsample\_genes} function, which resembles the \textit{Seurat} V3 VST gene selection method. All other parameters were left unmodified.  .
